---
title: "ifnb_walkthrough"
author: "Elias Williams"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: false
vignette: >
  %\VignetteIndexEntry{ifnb_walkthrough}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
require("devtools")
devtools::install_github("eliaswilliams/IntegrationScore", build_vignettes = FALSE)
library("IntegrationScore")
library(SeuratData)
library(Seurat)
library(ggplot2)
library(dplyr)
```

This is a vignette showing how to use IntegrationScore. In this vignette we 
use the ifnb scRNA-seq example data set sourced from the SeuratData package from 
the Satija Lab. (https://github.com/satijalab/seurat-data)

Integration steps taken from Seurat's vignette: 
https://satijalab.org/seurat/articles/integration_introduction.html

# Set Up

We will be using the interferon beta treated PBMC data set from Seurat Data. 


```{r load the data set, message=FALSE, warning=FALSE}
SeuratData::InstallData("ifnb")
```

## Spliting the Seurat Object into Two Batches and Clustering Independently

This data set consists of two group of PBMC cells. One group was treated with 
interferon beta (STIM group) the other was not (CTRL group). We begin our 
analysis by first splitting the Seurat object by the "stim" variable, which 
puts our cells into two groups: "STIM" and "CTRL". These will be our two batches. 

```{r, warning=FALSE, message = FALSE}
ifnb = subset(x = ifnb, downsample = 10000)
# split the dataset into a list of two seurat objects (stim and CTRL)
ifnb_split <- SplitObject(ifnb, split.by = "stim")
```

Now we will go through standard scRNA-seq pre-processing functions, culminating 
in clustering. It is important that we cluster each batch independently as we 
will be comparing each batch to itself in this analysis. 

```{r, warning=FALSE, message = FALSE}
# cluster each batch independently 
ifnb_split <- lapply(X = ifnb_split, FUN = function(x) {
        x <- NormalizeData(x)
        x <- FindVariableFeatures(x, selection.method="vst", nfeatures=3000)
        x <- ScaleData(x, features=VariableFeatures(object=x))
        x <- RunPCA(x, npcs=100, verbose = FALSE)
        x <- FindNeighbors(x, dims=1:20)
        x <- FindClusters(x, resolution = 0.3)
})

# make sure the cluster names are distinct between batches
levels(ifnb_split$CTRL@meta.data[["seurat_clusters"]]) = paste0(levels(ifnb_split$CTRL@meta.data[["seurat_clusters"]]), "_CTRL")
levels(ifnb_split$STIM@meta.data[["seurat_clusters"]]) = paste0(levels(ifnb_split$STIM@meta.data[["seurat_clusters"]]), "_STIM")
```

## Using Seurat's Built-in Integration Method to Batch Correct

Now we will run the built in batch correction method in Seurat v4. These steps 
were taken directly from their vignette. 
https://satijalab.org/seurat/articles/integration_introduction.html
```{r, warning=FALSE, message = FALSE}
# select features that are repeatedly variable across batches for integration
features <- SelectIntegrationFeatures(object.list = ifnb_split)

# find anchors for integration
anchors <- FindIntegrationAnchors(object.list = ifnb_split, anchor.features = features)

# run correction alg
ifnb_corrected <- IntegrateData(anchorset = anchors)

# make sure we are using right assay
DefaultAssay(ifnb_corrected) <- "integrated"
```

# Using IntegrationScore

## Plotting UMAPs

As a first step it is often useful to see how the batch correction has moved our 
data points around in some embedding space. UMAP is an excellent way to show our 
high dimensional gene expression data in two-dimensions. Here we run plotUMAP
to see a before and after look at our data.

```{r plotUMAP, warning=FALSE, fig.dim = c(7.5, 4)}
plotUMAP(ifnb_split, ifnb_corrected, colorBy="stim", nFeatures=500, nPC=10)
```

## Running findDEG

Now that we have an integration to score... let's score it!

First we run findDEG to find the top differentially expressed genes between 
clusters. We find the DEGs in both the corrected and uncorrected data sets. We 
will then use the results as a foundation of our evaluation. clusterVars 
indicated the clusters to use in each batch. logFCThresh is the cutoff log2FC.

```{r findDEG, warning=FALSE, message = FALSE}
deg_df <- findDEG(ifnb_split, ifnb_split_corrected, clusterVars = c('seurat_clusters','seurat_clusters'), logFCThresh=2)
head(deg_df)
```
Differentially expressed genes are measured in log2(fold change). For example, 
we may have an entry in our dataframe for CD4 in cluster 0 with a log2FC of 3. 
This would tell us that the gene expression for CD4 is 2**3 TIMES GREATER in 
cluster 0 compared to the average expression in other clusters. Cool! 

## Plotting DEG Deltas 

Now that we have our DEG dataframe, we are ready to see how Seurat's batch 
correction algorithm effects the gene expressions in our data set.

Here we run plotDelta to see how the change in logFC is distributed across 
batch. We also get to see some of the most changed genes and how their deltas 
are distrubuted across clusters.

```{r plotDelta, warning=FALSE, message=FALSE, fig.dim = c(7.5, 4)}
plotDelta(deg_df, 16) 
```

## Outlier Scores

Now that we've seen some delta plots, let's take a look at some correlations. 
plotCorrByCutoff gives us an easy way to see how the pre/post correction 
correlation between the differential expression of our genes changes as we 
consider more or less genes. The cutoff value in these plots can be interpreted 
as the number of top genes (ordered by the pre correction log2FC) we consider 
when calculating the correlation. Try setting byCluster to True to see an even 
more fine-grained look.

```{r plotCorrByCutoff, warning=FALSE, message=FALSE, fig.dim = c(7.5, 4)}
plotCorrByCutoff(deg_df, granularity = 100, estimate = "spearman", byCluster = T)
```

## Package References

[Williams, E. (2021) IntegrationScore: Comparison of Effectiveness of Batch Correction methods for  scRNA-seq For BCB410H. Unpublished.](https://github.com/eliaswilliams/IntegrationScore

<br>

## Other References

Büttner, M., Miao, Z., Wolf, F.A. et al. (2019) A test metric for assessing single-cell RNA-seq batch correction. Nat Methods 16, 43–49. https://doi.org/10.1038/s41592-018-0254-1

Hoffman, P. (2021) Satija Lab. Introduction to scRNA-seq integration. https://satijalab.org/seurat/articles/integration_introduction.html

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. https://www.R-project.org/

Satija, R. (2021) Satija Lab. Seurat Data R Package. https://github.com/satijalab/seurat-data

Wickham, H. and Bryan, J. (2019). R Packages (2nd edition). Newton, Massachusetts: O’Reilly Media. https://r-pkgs.org/


----

```{r}
sessionInfo()
```

